Diagnosising the sensitivity of grounding line flux to changes in sub-ice shelf melting

### 2019-07-08 draft_v2/draft_v2.tex
Updates from Overleaf

\begin{abstract}
Motivated by previous work using ice flow models to quantify the impacts of ice shelf buttressing on grounding line flux \citep[e.g.,][]{furst2016,reese2018}, we look for an improved physical understanding for how ice dynamics link ice thickness perturbations, via changes in sub-ice shelf melting, to changes in ice shelf buttressing and grounding line flux. More specifically, we seek to define one or more ``metrics'' that are 1) readily calculated from standard ice sheet model outputs and 2) informative with respect to diagnosing the sensitivity of grounding line flux to changes in ice thickness at specific locations on an ice shelf. By studying the ice dynamics for both idealized (MISMIP+) and realistic (Larsen C) ice shelves, we find a strongly direction-dependent buttressing number that links local changes in ice shelf thickness and dynamics with changes in the integrated grounding line flux. This buttressing metric, defined using the first principal stress, is better overall for quantifying changes in grounding line flux than a similar metric defined using the second principal stress or stress along the flow direction. This correlation is possibly controlled by the relative relationship of geometric/dynamic features between the perturbation point and the grounding line, indicating a dynamic (time evolving) sensitivity field of grounding line flux change to basal melt. Unfortunately, this metric only shows a robust relationship with the integrated grounding line flux for regions near the center of an ice shelf; for points too near the grounding line or the calving front, no clear relationship exists between any of the readily calculable metrics explored here and changes in grounding line flux. This motivates our exploration of an adjoint-based method for defining grounding line flux sensitivity to local changes in ice shelf geometry. When using the same idealized case, we demonstrate that this adjoint-based method is equivalent to the sensitivity analysis of \citep{reese2018} while requiring only a single model adjoint solve. We conclude that adjoint-based methods provide a powerful and efficient tool for determining the grounding line flux sensitivity to changes in sub-ice shelf melting.
\end{abstract}

\copyrightstatement{TEXT}

@@

Marine ice sheets like West Antarctica (and to a lesser extent, portions of East Antarctica) are grounded below sea level and their bedrock would remain so even after full isostatic rebound \citep{barletta2018}. This and the fact that ice sheets generally thicken inland lead to a geometric configuration that are probe to instability; a small increase in flux at the grounding line thins the ice there, leading to floatation, a retreat of the grounding line into deeper water, further increases in flux (due to still thicker ice), and further thinning and grounding line retreat. This theoretical ``marine ice sheet instability'' mechanism \citep{mercer1978, schoof2007} is supported by idealized \citep[e.g.,][]{schoof2012, asay2016} and realistic \citep[e.g.,][]{Cornford2015,royston2016} ice sheet modeling experiments and some studies \citep{joughin2014,rignot2014} argue that such an instability is currently under way along outlet glaciers of Antarctica's Amundsen Sea Embayment (ASE). The relevant perturbation for grounding line retreat in the ASE is thought to be intrusions of relatively warm, intermediate-depth ocean waters onto the continental shelves, which have reduced the thickness and extent of marginal ice shelves via increased submarine melting \citep[e.g.,][]{JenkinsEtAl2016}. These reductions are critical because fringing ice shelves restrict the flux of ice across their grounding lines farther upstream---the so-called ``buttressing'' affect of ice shelves \citep{gudmundsson2012, gudmundsson2013, derydt2015}---which makes them a critical control on the rate of ice flux from Antarctica to the ocean.

@@

Several recent studies apply whole-Antarctic ice sheet models, optimized to present-day observations, towards improving our understanding for how Antarctic ice shelves limit flux across the grounding line (and by extension how they impact ice dynamics farther inland). \citet{furst2016} calculated the buttressing across Antarctica ice shelves along two major directions (aligned with the ice flow and with the second principal stress) and evaluated their impacts on upstream ice dynamics to identify regions of the ice shelves that are dynamically ``passive''; in these regions increased submarine melting, or even complete removal of the ice should not significantly alter local or regional ice dynamics or the flux of from ice upstream. \citet{reese2018} used perturbation experiments to link small, localized decreases in ice shelf thickness to changes in integrated grounding line flux (GLF), thereby providing a map of GLF sensitivity to local increases in submarine melt rates. \citet{Goldberg2019} assessed the consistency of satellite retrieved and modeled melt rates, found the importance of bathymetry in influencing the ice sheet responses predicted by models and also investigated the adjoint sensitivity of volume above floatation to melt rates for Dotson and Crosson ice shelves, Antarctica.

@@

\textbf{Steve to update this paragraph to better reflect what we actually do below: 1) connect the studies of Furst and Reese by trying to understand if / how local evaluations of ice shelf buttressing do or do not reflect the actual potential for purturbations at those same locations to impact the flux across the GL (since that is, presumably, why we care about buttressing in the first place -- noting that what we'd like is some easily calculated 'metric' to identify these important locations for us); 2) assess the ability of analagous, relatively easy to calculate, local buttressing metrics to indicate GLF sensitivity to local perturbations in ice shelf geometry; 3) point out where these metrics are useful and where (and why) they are not; 4) propose and demonstrate alternate, adjoint-based method that has advantages over previous methods.}
Motivated by these studies, we build on and extend the methods and analysis of \citet{furst2016} and \citet{reese2018} in order to make progress towards answering the following questions: 
(1) How do local and regional changes in ice shelf geometry affect distal changes in GLF? 
%(2) Can local or regional ice shelf dynamics explain GLF sensitivity to local or regional changes in ice shelf thickness? 
(2) Can we derive and define new analyses for understanding how observed or modeled spatial patterns in submarine melting influence GLF and, by extension, project how changes in submarine melt pattern and magnitude will impact GLF in the future?

@@

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{figs/mismip+_larsenC.pdf}
\caption{Plan view of steady-state surface speed for a) MISMIP+ and b) Larsen C Ice Shelf. The black curves show the grounding lines.}
\label{mismip+_larsenC}
\end{figure}

\section{Model description}
We use the MPAS-Albany Land Ice model \citep[MALI;][]{hoffman2018}, which solves the three-dimensional, first-order approximation to the Stokes momentum balance for ice flow\footnote{See \citet{schoof2013} for a full description of the Stokes momentum balance for ice flow and its lower-order approximations.}. Using the notation  of \citet{perego2012} and \citet{tezaur2015a} this can be expressed as, 
\begin{equation} \label{eq:foStokes}
\left\{
\begin{array}{rcl} -\nabla \cdot (2 \mu_e \dot{\boldsymbol{\epsilon}}_1) + \rho_{i} g
\frac{\partial s}{\partial x}&=&0, \\
-\nabla \cdot (2 \mu_e \dot{\boldsymbol{\epsilon}}_2) +\rho_{i} g
\frac{\partial s}{\partial y} &=& 0, \\
\end{array}\right.
\end{equation}
where $x$ and $y$ are the horizontal coordinate vectors in a Cartesian reference frame, $s(x,y)$ is the ice surface elevation, $\rho_{i}$ represents the ice density, $g$ the acceleration due to gravity, and $\dot{\boldsymbol{\epsilon}}_{1,2}$ are the two dimensional strain rate vectors given by
\begin{equation}
\dot{\boldsymbol{\epsilon}}_1 = \left(\begin{array}{ccc}
2\dot{\epsilon}_{xx} + \dot{\epsilon}_{yy}, &\dot{\epsilon}_{xy},&
\dot{\epsilon}_{xz}\end{array}\right)^T,
\end{equation}
and
\begin{equation}
\dot{\boldsymbol{\epsilon}}_2 = \left(
\begin{array}{ccc}\dot{\epsilon}_{xy}, &
\dot{\epsilon}_{xx} + 2\dot{\epsilon}_{yy}, &\dot{\epsilon}_{yz}
\end{array}\right)^T.
\end{equation}
The ``effective'' ice viscosity, $ \mu_e$ in  (\ref{eq:foStokes}), is given by 
\begin{equation}
\label{eq:effvisc}
    \mu_{e}~=~\gamma A^{-\frac{1}{n}}\dot{\epsilon}_{e}^{\frac{1-n}{n}},
\end{equation}
where $\gamma$ is an ice stiffness factor, $A$ is a temperature-dependent rate factor, $n=3$ is the power-law exponent, and the effective strain rate, $\dot{\epsilon}_{e}$, is defined as
\begin{equation} \label{eq:effstrain}
\dot{\epsilon}_e \equiv \left( \dot{\epsilon}_{xx}^2 +
\dot{\epsilon}_{yy}^2 + \dot{\epsilon}_{xx} \dot{\epsilon}_{yy} +
\dot{\epsilon}_{xy}^2 + \dot{\epsilon}_{xz}^2 +
\dot{\epsilon}_{yz}^2 \right)^\frac{1}{2}.
\end{equation}
Gradients in the horizontal velocity components, $u$ and $v$, contribute to the individual strain rate terms in (\ref{eq:effstrain}) and are given by
\begin{eqnarray} \label{eq:epsilonij}
\dot{\epsilon}_{xx} &=& \frac{\partial u}{\partial x}, \hspace{0.5cm} 
\dot{\epsilon}_{yy} = \frac{\partial v}{\partial y}, \hspace{0.5cm}
\dot{\epsilon}_{xy} = \frac{1}{2}\left(\frac{\partial u}{\partial y} + \frac{\partial v}{\partial x} \right), \nonumber\\ 
\dot{\epsilon}_{xz} &=& \frac{1}{2} \frac{\partial u}{\partial z}, \textrm{ and } \hspace{0.15cm} 
\dot{\epsilon}_{yz} = \frac{1}{2} \frac{\partial v}{\partial z}.
\end{eqnarray}

A stress free upper surface is enforced through 
\begin{equation} \label{eq:stressFreeBC}
\dot{\boldsymbol{\epsilon}}_1 \cdot \mathbf{n} = \dot{\boldsymbol{\epsilon}}_2 \cdot \mathbf{n} = 0, 
\end{equation}
%\begin{equation} \label{eq:stressFreeBC}
%\boldsymbol{\sigma} \cdot \mathbf{n} = 0, 
%\end{equation}
where $\mathbf{n}$ is the outward pointing normal vector at the ice sheet upper surface, $z=s(x,y)$.
The lower surface is allowed to slide according to the continuity of basal tractions,
\begin{equation} \label{eq:basalbc}
\begin{array}{ll}
2\mu_e \dot{\boldsymbol{\epsilon}}_1 \cdot \mathbf{n} + \beta u = 0, \hspace{0.2cm} 2 \mu \dot{\boldsymbol{\epsilon}}_2 \cdot \mathbf{n} + \beta v = 0, \\
\end{array}
\end{equation}
%\begin{equation} \label{eq:basalbc}
%\boldsymbol{\sigma} \cdot \mathbf{n} + \beta \boldsymbol{u} = 0,
%\end{equation}
where $\beta$ is a spatially variable friction coefficient, $\boldsymbol{\sigma}$ is the stress tensor and $\boldsymbol{u}$ is the two-dimensional velocity vector ($u$, $v$). On lateral boundaries in contact with the ocean, the portion of the boundary above sea level is stress free while the portion below sea level feels the ocean hydrostatic pressure according to
\begin{equation}\label{eq:oceanbc}
\begin{array}{ll}
2 \mu_e \dot{\boldsymbol{\epsilon}}_1 \cdot \mathbf{n} = \frac{1}{2} \rho_i g H \left(1- \frac{\rho_i}{\rho_o} \right) \mathbf{n}_1, \quad 2 \mu_e  \dot{\boldsymbol{\epsilon}}_2 \cdot \mathbf{n} = \frac{1}{2} \rho_i g H \left(1- \frac{\rho_i}{\rho_o} \right) \mathbf{n}_2
\end{array}
\end{equation}
%\begin{equation}\label{eq:oceanbc}
%\boldsymbol{\sigma} \cdot \mathbf{n} = \rho_i g (s-z)\mathbf{n},
%\end{equation}
%where $\rho_o$ represents the density of ocean water and $xx\mathbf{n}$ the outward pointing normal vector to the lateral boundary (i.e., parallel to the $(x,y)$ plane).
where $\mathbf{n}$ the outward pointing normal vector to the lateral boundary (i.e., parallel to the $(x,y)$ plane).
%\textbf{TZ: I change this section a bit to avoid it looking identical to Matt's MALI paper.}

@@

Here, we apply MALI to experiments on both idealized and realistic marine-ice sheet geometries. For our idealized domain and model state, we start from the equilibrium initial conditions for the MISMIP+ experiments, as described in \citet{asay2016}. The model mesh is spatially uniform at 2 km resolution. For our realistic domain, we use Antarctica's Larsen C Ice Shelf and its upstream catchment area. The model state is based on the optimization of the ice stiffness ($\gamma$ in Eqn (\ref{eq:effvisc})) and basal friction ($\beta$ in Eqn (\ref{eq:basalbc})) coefficients in order to provide a best match between modeled and observed present-day velocities \citep{rignot2014} using adjoint-based methods discussed in \citet{perego2014} and \citet{hoffman2018}. The domain geometry is based on Bedmap2 \citep{fretwell2013} and ice temperatures, which are held fixed for this study, are based on \citet{liefferinge2013}. Mesh resolution on the ice shelf is between 2 and 6 km and coarsens to 20 km in the ice sheet interior. Following optimization to present-day velocities, the model is relaxed using a 100-year forward run, providing the initial condition from which the Larsen C experiments are conducted (as discussed below). Both the MISMIP+ and Larsen C experiments use 10 vertical layers that are finest near the bed and coarsen towards the surface. The grounding line position is determined from hydrostatic equilibrium. A sub-element parameterization method is used to define basal friction coefficient values at the grounding line \citep{seroussi2014} .

@@

To explore the sensitivity of changes in GLF to small, localized changes in ice shelf thickness, we conduct a number of perturbation experiments analogous to those of \citet{reese2018}. Using diagnostic (instantaneous) model solutions, we first study the instantaneous response of GLF for the idealized geometry and initial state provided by the MISMIP+ experiment \citep{asay2016}. We then conduct a similar study for Antarctica's Larsen C Ice Shelf using a realistic configuration and initial state.

Our experiments are conducted in a manner similar to those of \citet{reese2018}. We perturb the coupled ice sheet-shelf system by decreasing the ice thickness uniformly by 1 m over square grid ``boxes'' covering the base of the ice shelves, after which we examine the instantaneous impact on kinematics and dynamics (discussed further below). For MISMIP+, we use a uniform, hexagon-type, around 2-km mesh (for simplicity we use the term 2-km in the following). For the Larsen C Ice Shelf, horizontal mesh resolution is spatially variable and we assign each grid cell to fall within one and only one box based on its location. For MISMIP+, we perturb the thickness of the cell (hexagon) in the mesh. For Larsen C, we only use 20$\times$20 km square boxes \citep[the same resolution as in][]{reese2018}. Lastly, for the MISMIP+ 2-km experiments we note that, in order to save on computing costs, we only perturb the region of the ice shelf for which $x<530$ km (the area over which the ice shelf is laterally buttressed).  We also perturb only half of the ice shelf, $y>40$ km, due to the symmetry about the center line, though we analyze the response to these perturbations over the entire model domain.

>>>>

\begin{abstract}
In this paper, we look for an improved physical understanding for how ice dynamics link ice thickness perturbations, via changes in sub-ice shelf melting, to changes in ice shelf buttressing and grounding line flux. More specifically, we seek a connection among grounding line flux, ice shelf buttressing and basal melting. By studying the ice dynamics for both idealized (MISMIP+) and realistic (Larsen C) ice shelves, we find a strongly direction-dependent buttressing number that links local changes in ice shelf thickness and dynamics with changes in the integrated grounding line flux. From these two examples, this buttressing metric, defined using the first principal stress, is better overall for quantifying changes in grounding line flux than a similar metric defined using the second principal stress or stress along the flow direction. This correlation is possibly controlled by the relative relationship of geometric/dynamic features between the perturbation point and the grounding line, indicating a dynamic (time evolving) sensitivity field of integrated grounding line flux to basal melt. We find this buttressing metric only shows a robust relationship with the integrated grounding line flux for regions near the center of an ice shelf; for points too near the grounding line or the calving front, no clear relationship exists. This motivates our exploration of an adjoint-based method for defining integrated grounding line flux sensitivity to local changes in ice shelf geometry. \textbf{TZ: add 1-2 sentences describing the results of adjoint when we get the larsen C results, for example, what does adjoint look like near GL?}
\end{abstract}

%\copyrightstatement{TEXT}

@@

Marine ice sheets like West Antarctica (and to a lesser extent, portions of East Antarctica) are grounded below sea level and their bedrock would remain so even after full isostatic rebound \citep{barletta2018}. This and the fact that ice sheets generally thicken inland lead to a geometric configuration that are prone to instability; a small increase in flux at the grounding line thins the ice there, leading to floatation, a retreat of the grounding line into deeper water, further increases in flux (due to still thicker ice), and further thinning and grounding line retreat. This theoretical ``marine ice sheet instability'' mechanism \citep{mercer1978, schoof2007} is supported by idealized \citep[e.g.,][]{schoof2012, asay2016} and realistic \citep[e.g.,][]{Cornford2015,royston2016} ice sheet modeling experiments and some studies \citep{joughin2014,rignot2014} argue that such an instability is currently under way along outlet glaciers of Antarctica's Amundsen Sea Embayment (ASE). The relevant perturbation for grounding line retreat in the ASE is thought to be intrusions of relatively warm, intermediate-depth ocean waters onto the continental shelves, which have reduced the thickness and extent of marginal ice shelves via increased submarine melting \citep[e.g.,][]{JenkinsEtAl2016}. These reductions are critical because fringing ice shelves restrict the flux of ice across their grounding lines farther upstream---the so-called ``buttressing'' affect of ice shelves \citep{gudmundsson2012, gudmundsson2013, derydt2015}---which makes them a critical control on the rate of ice flux from Antarctica to the ocean.

@@

Several recent studies apply whole-Antarctic ice sheet models, optimized to present-day observations, towards improving our understanding for how Antarctic ice shelves limit flux across the grounding line (and by extension how they impact ice dynamics farther inland). \citet{furst2016} calculated the buttressing number across Antarctica ice shelves along two major directions (aligned with the ice flow and with the second principal stress) and evaluated their impacts on upstream ice dynamics to identify regions of the ice shelves that are dynamically ``passive''; in these regions increased submarine melting, or even complete removal of the ice should not significantly alter local or regional ice dynamics or the flux of from ice upstream. \citet{reese2018} used perturbation experiments to link small, localized decreases in ice shelf thickness to changes in integrated grounding line flux (GLF), thereby providing a map of GLF sensitivity to local increases in submarine melt rates. \citet{Goldberg2019} assessed the consistency of satellite retrieved and modeled melt rates, found the importance of bathymetry in influencing the ice sheet responses predicted by models and also investigated the adjoint sensitivity of volume above floatation to melt rates for Dotson and Crosson ice shelves, Antarctica.

@@

\textbf{Steve to update this paragraph to better reflect what we actually do below: 1) connect the studies of Furst and Reese by trying to understand if / how local evaluations of ice shelf buttressing do or do not reflect the actual potential for purturbations at those same locations to impact the flux across the GL (since that is, presumably, why we care about buttressing in the first place -- noting that what we'd like is some easily calculated 'metric' to identify these important locations for us); 2) assess the ability of analagous, relatively easy to calculate, local buttressing metrics to indicate GLF sensitivity to local perturbations in ice shelf geometry; 3) point out where these metrics are useful and where (and why) they are not; 4) propose and demonstrate alternate, adjoint-based method that has advantages over previous methods.}
Motivated by these studies, we build on and extend the methods and analysis of \citet{furst2016} and \citet{reese2018} in order to make progress towards answering the following questions: 
(1) Is there any easily accessible metrics that indicate the sensitivity of GLF to the thickness change at certain locations on ice shelf?
(2) Is buttressing number a such metric? Does thickness decrease at the locations with large buttressing number results in large GLF increase?
(3) What are the mechanisms that connect ice thickness perturbation (basal melting) and GLF changes?
%(1) How do local and regional changes in ice shelf geometry affect distal changes in GLF? 
%(2) Can local or regional ice shelf dynamics explain GLF sensitivity to local or regional changes in ice shelf thickness? 
(4) Can we derive easy-to-implement and new methods/metrics for indicating the GLF sensitivity of basal melting across ice shelf during the model run over time so that we can better project how changes in submarine melt pattern and magnitude will impact GLF in the future?

@@

\begin{figure}
\centering
\includegraphics[width=1\linewidth]{./figs/mismip_larsenc.pdf}
\caption{Plan view of steady-state surface speed for a) MISMIP+ and b) Larsen C Ice Shelf. The black curves show the grounding lines.}
\label{mismip_larsenc}
\end{figure}

\section{Model description}
We use the MPAS-Albany Land Ice model \citep[MALI;][]{hoffman2018}, which solves the three-dimensional, first-order approximation to the Stokes momentum balance for ice flow\footnote{See \citet{schoof2013} for a full description of the Stokes momentum balance for ice flow and its lower-order approximations.}. Using the notation  of \citet{perego2012} and \citet{tezaur2015a} this can be expressed as, 
\begin{equation} \label{eq:foStokes}
\left\{
\begin{array}{rcl} -\nabla \cdot (2 \mu_e \dot{\boldsymbol{\epsilon}}_1) + \rho_{i} g
\frac{\partial s}{\partial x}&=&0, \\
-\nabla \cdot (2 \mu_e \dot{\boldsymbol{\epsilon}}_2) +\rho_{i} g
\frac{\partial s}{\partial y} &=& 0, \\
\end{array}\right.
\end{equation}
where $x$ and $y$ are the horizontal coordinate vectors in a Cartesian reference frame, $s(x,y)$ is the ice surface elevation, $\rho_{i}$ represents the ice density, $g$ the acceleration due to gravity, and $\dot{\boldsymbol{\epsilon}}_{1,2}$ are the two dimensional strain rate vectors given by
\begin{equation}
\dot{\boldsymbol{\epsilon}}_1 = \left(\begin{array}{ccc}
2\dot{\epsilon}_{xx} + \dot{\epsilon}_{yy}, &\dot{\epsilon}_{xy},&
\dot{\epsilon}_{xz}\end{array}\right)^T,
\end{equation}
and
\begin{equation}
\dot{\boldsymbol{\epsilon}}_2 = \left(
\begin{array}{ccc}\dot{\epsilon}_{xy}, &
\dot{\epsilon}_{xx} + 2\dot{\epsilon}_{yy}, &\dot{\epsilon}_{yz}
\end{array}\right)^T.
\end{equation}
The ``effective'' ice viscosity, $ \mu_e$ in  Eq.~(\ref{eq:foStokes}), is given by 
\begin{equation}
\label{eq:effvisc}
    \mu_{e}~=~\gamma A^{-\frac{1}{n}}\dot{\epsilon}_{e}^{\frac{1-n}{n}},
\end{equation}
where $\gamma$ is an ice stiffness factor, $A$ is a temperature-dependent rate factor, $n=3$ is the power-law exponent, and the effective strain rate, $\dot{\epsilon}_{e}$, is defined as
\begin{equation} \label{eq:effstrain}
\dot{\epsilon}_e \equiv \left( \dot{\epsilon}_{xx}^2 +
\dot{\epsilon}_{yy}^2 + \dot{\epsilon}_{xx} \dot{\epsilon}_{yy} +
\dot{\epsilon}_{xy}^2 + \dot{\epsilon}_{xz}^2 +
\dot{\epsilon}_{yz}^2 \right)^\frac{1}{2},
\end{equation}
where $\dot{\epsilon}_{ij}$ are the corresponding strain-rate components.
%Gradients in the horizontal velocity components, $u$ and $v$, contribute to the individual strain rate terms in (\ref{eq:effstrain}) and are given by
%\begin{eqnarray} \label{eq:epsilonij}
%\dot{\epsilon}_{xx} &=& \frac{\partial u}{\partial x}, \hspace{0.5cm} 
%\dot{\epsilon}_{yy} = \frac{\partial v}{\partial y}, \hspace{0.5cm}
%\dot{\epsilon}_{xy} = \frac{1}{2}\left(\frac{\partial u}{\partial y} + \frac{\partial v}{\partial x} \right), \nonumber\\ 
%\dot{\epsilon}_{xz} &=& \frac{1}{2} \frac{\partial u}{\partial z}, \textrm{ and } \hspace{0.15cm} 
%\dot{\epsilon}_{yz} = \frac{1}{2} \frac{\partial v}{\partial z}.
%\end{eqnarray}

A stress free upper surface is enforced through 
\begin{equation} \label{eq:stressFreeBC}
\dot{\boldsymbol{\epsilon}}_1 \cdot \mathbf{n} = \dot{\boldsymbol{\epsilon}}_2 \cdot \mathbf{n} = 0, 
\end{equation}
%\begin{equation} \label{eq:stressFreeBC}
%\boldsymbol{\sigma} \cdot \mathbf{n} = 0, 
%\end{equation}
where $\mathbf{n}$ is the outward pointing normal vector at the ice sheet upper surface, $z=s(x,y)$.
The lower surface is allowed to slide according to the continuity of basal tractions,
\begin{equation} \label{eq:basalbc}
\begin{array}{ll}
2\mu_e \dot{\boldsymbol{\epsilon}}_1 \cdot \mathbf{n} + \beta u = 0, \hspace{0.2cm} 2 \mu \dot{\boldsymbol{\epsilon}}_2 \cdot \mathbf{n} + \beta v = 0, \\
\end{array}
\end{equation}
%\begin{equation} \label{eq:basalbc}
%\boldsymbol{\sigma} \cdot \mathbf{n} + \beta \boldsymbol{u} = 0,
%\end{equation}
where $\beta$ is a spatially variable friction coefficient, $\boldsymbol{\sigma}$ is the stress tensor and $\boldsymbol{u}$ is the two-dimensional velocity vector ($u$, $v$). On lateral boundaries in contact with the ocean, the portion of the boundary above sea level is stress free while the portion below sea level feels the ocean hydrostatic pressure according to
\begin{equation}\label{eq:oceanbc}
\begin{array}{ll}
2 \mu_e \dot{\boldsymbol{\epsilon}}_1 \cdot \mathbf{n} = \frac{1}{2} \rho_i g H \left(1- \frac{\rho_i}{\rho_o} \right) \mathbf{n}_1, \quad 2 \mu_e  \dot{\boldsymbol{\epsilon}}_2 \cdot \mathbf{n} = \frac{1}{2} \rho_i g H \left(1- \frac{\rho_i}{\rho_o} \right) \mathbf{n}_2
\end{array}
\end{equation}
%\begin{equation}\label{eq:oceanbc}
%\boldsymbol{\sigma} \cdot \mathbf{n} = \rho_i g (s-z)\mathbf{n},
%\end{equation}
%where $\rho_o$ represents the density of ocean water and $xx\mathbf{n}$ the outward pointing normal vector to the lateral boundary (i.e., parallel to the $(x,y)$ plane).
where $\mathbf{n}$ is the outward pointing normal vector to the lateral boundary (i.e., parallel to the $(x,y)$ plane) and $\rho_o$ is the density of ocean water.
%\textbf{TZ: I change this section a bit to avoid it looking identical to Matt's MALI paper.}

@@

Here, we apply MALI to experiments on both idealized and realistic marine-ice sheet geometries. For our idealized domain and model state, we start from the equilibrium initial conditions for the MISMIP+ experiments, as described in \citet{asay2016}. The model mesh is spatially uniform at an around 2 km resolution. For our realistic domain, we use Antarctica's Larsen C Ice Shelf and its upstream catchment area. The model state is based on the optimization of the ice stiffness ($\gamma$ in Eq. (\ref{eq:effvisc})) and basal friction ($\beta$ in Eq. (\ref{eq:basalbc})) coefficients in order to provide a best match between modeled and observed present-day velocities \citep{rignot2014} using adjoint-based methods discussed in \citet{perego2014} and \citet{hoffman2018}. The domain geometry is based on Bedmap2 \citep{fretwell2013} and ice temperatures, which are held fixed for this study, are based on \citet{liefferinge2013}. Mesh resolution on the ice shelf is between 2 and 6 km and coarsens to 20 km in the ice sheet interior. Following optimization to present-day velocities, the model is relaxed using a 100-year forward run, providing the initial condition from which the Larsen C experiments are conducted (as discussed below). Both the MISMIP+ and Larsen C experiments use 10 vertical layers that are finest near the bed and coarsen towards the surface. The grounding line position is determined from hydrostatic equilibrium. A sub-element parameterization method is used to define basal friction coefficient values at the grounding line \citep{seroussi2014}.

@@

To explore the sensitivity of changes in GLF to small, localized changes in ice shelf thickness, we conduct a number of perturbation experiments analogous to those of \citet{reese2018}. Using diagnostic (instantaneous) model solutions, we first study the instantaneous response of GLF for the idealized geometry and initial state provided by the MISMIP+ experiment \citep{asay2016}. We then conduct a similar study for Antarctica's Larsen C Ice Shelf using a realistic configuration and initial state. The geometry and steady-state ice speed maps for MISMIP+ and Larsen C ice shelf are shown in Fig.~\ref{mismip_larsenc}.

Our experiments are conducted in a manner similar to those of \citet{reese2018}. We perturb the coupled ice sheet-shelf system by decreasing the ice thickness uniformly by 1 m over square grid ``boxes'' covering the base of the ice shelves, after which we examine the instantaneous impact on kinematics and dynamics (discussed further below). For MISMIP+, we use a uniform, hexagon-type, around 2-km mesh (for simplicity we use the term 2-km in the following). For the Larsen C Ice Shelf, horizontal mesh resolution is spatially variable and we assign each grid cell to fall within one and only one box based on its location. For MISMIP+, we perturb the thickness of the cell (hexagon) in the mesh. For Larsen C, we only use 20$\times$20 km square boxes \citep[the same resolution as in][]{reese2018}. Lastly, for the MISMIP+ 2-km experiments we note that, in order to save on computing costs, we only perturb the region of the ice shelf for which $x<530$ km (the area over which the ice shelf is laterally buttressed).  We also perturb only half of the ice shelf, $y>40$ km, due to the symmetry about the center line, though we analyze the response to these perturbations over the entire model domain. \textbf{TZ: I am afraid I still want to use Python to make Fig. 1 to show the coordinates}
